Presentation Title

### 2017-08-22 thesis/chapters/ch1-intro.tex
Edit grammar etc.

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}, 
which is due to the fact that blocks are fixed to 1 MB and are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
A few months ago the backlog even reached 100,000, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some call a civil war~\cite{bitcoincivilwar}.
Parties which are for an increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
It also requires a hard (non-backwards compatible) fork,
which risks consensus failure and devaluation.
A recent empirical study by Croman et al.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
But given the bandwidth and latency constraints of today's network,
it is not possible to have more than 27 transactions per second.
To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
Thus, fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
which is in the order of tens of thousands transactions per second~\cite{visa}.

>>>>

Sadly, as traditional blockchain systems began to gain popularity,
their limitations also became apparent.
Bitcoin has the infamous 7 transactions per second upper bound~\cite{vukolic2015quest}, 
which is due to the fact that blocks are fixed to 1 MB and are only generated on average every 10 minutes.
Since every Bitcoin transaction is at least 250 bytes, it computes to about 7 transactions per second.
Due to this limitation, it is not uncommon to see a long backlog of about 20,000 transactions\footnote{\url{https://blockchain.info/unconfirmed-transactions/}}.
The backlog even reached 100,000 in May 2017, which meant new transactions would take at least 11 hours to be written to the blockchain~\cite{bitcoinbacklog}.
This issue has plagued the Bitcoin community for some time and is the root cause for the block size debate, which some call a civil war~\cite{bitcoincivilwar}.
Parties which are for an increase in block size argue that a larger block would improve the transaction rate.
Parties against it argue that it would make mining more centralised because blocks take longer to propagate through the network.
It also requires a hard (non-backwards compatible) fork,
which risks consensus failure and devaluation.
A recent empirical study by Croman et al.~\cite{croman2016scaling} has shown that increasing the block size or the block interval may help.
But given the bandwidth and latency constraints of today's network,
it is not possible to have more than 27 transactions per second.
To achieve transaction rates higher than that, it will lead to unfairness for nodes that do not have a good network connection.
Thus, fundamental changes are necessary to run at the scale of centralised payment processors such as Visa,
which is in the order of tens of thousands transactions per second~\cite{visa}.

### 2017-08-22 thesis/chapters/ch2-problem.tex
Edit grammar etc.

Blockchain research has seen a surge in recent years from both the industry and academia,
this resulted in a myriad of blockchain systems with different approaches to scalability.
To understand the current state-of-the-art,
we classify various block\-chain systems by their scalability approach.

@@

Off-chain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and hashed time-locked Bitcoin contracts~\cite{bitcointimelock}.
In the simplest case, 
if two parties wish to make frequent transactions,
they open a payment channel in the form of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
Each of the party must also deposit some Bitcoins in the multi-signature address,
this is called the opening transaction.
Both parties keep the channel state which is not broadcasted to the network.
The state is updated when the two parties make new transactions.
The protocol disincentivises the parties from broadcasting old channel states.
If this occurs, the counterparty can drain all the Bitcoins in the channel.
To close the channel, the parties simply broadcast the latest net settlement to the Bitcoin network,
this is called the closing transaction.
Effectively, only two transactions need to be broadcasted to the Bitcoin network---the opening and the closing transaction,
even if the two parties made thousands or millions of transactions in between.

@@

\subsection{Combining PoW with Byzantine consensus}
Recent research has developed a class of hybrid systems which uses PoW for committee election,
and Byzantine consensus algorithms to agree on transactions.
% And then they use a Byzantine consensus algorithm to actually reach consensus on a set of transactions within the committee.
Some examples are SCP~\cite{luu2015scp}, ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.

This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g., PBFT in ByzCoin~\cite{kogias2016enhancing})
while keeping the permissionless nature by using PoW.
A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
For SCP, ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
Classical blockchain does not have this problem because they do not use Byzantine consensus.
Further, due to the fact that these systems must reach consensus on all transactions, none of them achieves horizontal scalability.

@@

Sharding is a technique to achieve horizontal scalability by grouping nodes into multiple committees, also known as shards.
Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
An intra-shard protocol is needed for transactions that involve nodes from more than one shard.
The number of shards grows linearly with respect to the total computational power in the network.
This scheme achieves horizontal scalability because if every shard commits transactions at the same throughput,
then adding more shard would naturally result in a linear increase of global throughput.
Examples of blockchain systems that use sharding are Elastico~\cite{luu2016elastico} and Omni\-Ledger~\cite{kokoris2017omniledger}.

@@

We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
To the best of our knowledge, only the two systems described in~\Cref{sec:sharding} have experimentally demonstrated horizontal scalability.
However, these techniques rely on sharding, thus the performance highly depends on the transaction characteristics and the shard size parameter.
Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
We wish to design a blockchain consensus protocol that can be used in many applications.
For example, the accounting of detailed internet traffic in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market in Tribler\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
Hence our research question is the following.
\begin{displayquote}
\emph{How can we design a blockchain consensus protocol that is fault-tolerant,
horizontally scalable,
and able to reach global consensus?}
\end{displayquote}

In order to make sense of our research question, we must first explain \emph{blockchain consensus protocol}.
Typical blockchain systems are application specific.
For example, the application in Bitcoin (any many of its derivatives) is digital cash, the application in Namecoin is domain name registration~\cite{namecoin} and the application in Siacoin is cloud storage~\cite{siacoin}.
Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
In the case of Bitcoin, it is PoW,
because it is not concerned with the structure of the transactions.
In this work, we focus on the blockchain consensus protocol rather than any specific application because it is the scalability bottleneck and is necessary for any blockchain based application.
% Also, it should be possible to build any kind of application on top our system.
% We do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.

In the remainder of this section, we expand on our research question and visit each aspect in detail.
At the end of this section, we describe some issues that we do not consider as a part of our design,
which is the result of not focusing on any specific application.

@@

\subsection*{Fault-tolerance}
% WHY WE NEED FAULT-TOLERANCE
Similar to many blockchain systems,
our system should be unaffected in the presence of powerful adversaries.
In particular, adversaries are Byzantine,
meaning that they can have arbitrary behaviour.
Thus anything is possible from simply omitting messages to colluding with each other in order to undermine the whole system.

There is another aspect of fault tolerance that is transaction verifiability.
All transactions must be eventually verifiable to maintain system integrity.
Further, the validation result must be consistent.
For example, if two different honest nodes attempt to verify the same transaction,
it cannot be the case that one node thinks that the transaction is valid but the other thinks it is invalid.

@@

\subsection*{Limitations}
Since our consensus protocol is application neutral, it does not attempt to prevent the Sybil attack~\cite{douceur2002sybil}.
A good defence mechanism requires application specific information.
For example, a social network based Sybil defence mechanisms use random walks in real-world social graphs~\cite{yu2006sybilguard}.
Online marketplaces such as Amazon use the rating of buyer and sellers.
Cryptocurrencies use PoW to vote on the latest block.
% Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
For the same reason, we do not explicitly consider spam or denial of service attacks 
because without a concrete application,
it is difficult to distinguish a super active user from an attacker.
For this work,
we assume that it is the role of application developer to find a reputation mechanism or a trust model suitable for defending against such attacks.

>>>>

Blockchain research has seen a surge in recent years from both the industry and academia,
this resulted in a myriad of blockchain systems with different approaches to scalability.
To understand the current state-of-the-art,
we classify various block\-chain systems by their scalability approach.
A summary is given in~\Cref{tab:scalability}.

\begin{table}[htb]
    \centering
    \caption{Summary of the scalability properties of many blockchain systems.
    Scalability gets better from left to right.}
    \label{tab:scalability}
    \begin{tabular}{|l|l|l|l|}
    \hline
    \textbf{\thead{Not\\scalable}} & \textbf{\thead{Somewhat\\scalable}} & \textbf{\thead{Limited horizontal\\scalability}} & \textbf{\thead{Full horizontal\\scalability}} \\ \hline
    Bitcoin~\cite{bitcoin}           & Hyperledger~\cite{cachin2016architecture} & Elastico~\cite{luu2016elastico}           & \textsc{Checo} (this work)        \\
    Ethereum~\cite{wood2014ethereum} & ByzCoin~\cite{kogias2016enhancing}       & OmniLedger~\cite{kokoris2017omniledger}   &                                   \\
    etc.                             & Solidius~\cite{abraham2016solidus}       &                                           &                                   \\
    \hline
    \end{tabular}
\end{table}

@@

Off-chain transaction systems are implemented using multi-signature addresses~\cite{bitcoinmultisig} and hashed time-locked Bitcoin contracts~\cite{bitcointimelock}.
In the simplest case, 
if two parties wish to make frequent transactions,
they open a payment channel in the form of a multi-signature Bitcoin address (for two parties it would be a 2-of-2 signature address).
Each of the party must also deposit some Bitcoins in the multi-signature address,
which is called the opening transaction.
Both parties keep the channel state which is not broadcasted to the network.
The state is updated when the two parties make new transactions.
The protocol disincentivises the parties from broadcasting old channel states.
If this occurs, the counterparty can drain all the Bitcoins in the channel.
To close the channel, the parties simply broadcast the latest net settlement to the Bitcoin network,
which is called the closing transaction.
Effectively, only two transactions need to be broadcasted to the Bitcoin network---the opening and the closing transaction,
even if the two parties made thousands or millions of transactions in between.

@@

\subsection{Combining PoW with Byzantine consensus}
Recent research has developed a class of hybrid systems which uses PoW for committee election,
and Byzantine consensus algorithms to agree on transactions.
% And then they use a Byzantine consensus algorithm to actually reach consensus on a set of transactions within the committee.
Some examples are ByzCoin~\cite{kogias2016enhancing} and Solidus~\cite{abraham2016solidus}.

This technique overcomes the early blockchain scalability issue by delegating the transaction validation to the Byzantine consensus protocol (e.g., PBFT in ByzCoin~\cite{kogias2016enhancing})
while keeping the permissionless nature by using PoW.
A major tradeoff of such systems is that they cannot guarantee correctness when there is a large number of malicious nodes (but less than a majority).
For ByzCoin and Solidus, they all have some probability to elect more than $t$ Byzantine nodes into the committee,
where $t$ is typically just under a third of the committee size (a lower bound of Byzantine consensus~\cite{pease1980reaching}).
This problem is especially difficult to solve because the committee is always much smaller than the population size which has more than $t$ Byzantine nodes.
Classical blockchain does not have this problem because they do not use Byzantine consensus.
Further, due to the fact that these systems must reach consensus on all transactions, none of them achieves horizontal scalability.

@@

Sharding is a technique to achieve horizontal scalability by grouping nodes into multiple committees, also known as shards.
Nodes within a single shard run a Byzantine consensus algorithm to agree on a set of transactions that belong to that specific shard.
An inter-shard protocol is needed for transactions that involve nodes from more than one shard.
The number of shards grows linearly with respect to the total computational power in the network.
This scheme achieves horizontal scalability because if every shard commits transactions at the same throughput,
then adding more shard would naturally result in a linear increase of global throughput.
Examples of blockchain systems that use sharding are Elastico~\cite{luu2016elastico} and Omni\-Ledger~\cite{kokoris2017omniledger}.

@@

We saw in~\Cref{sec:taxonomy} that the majority of the systems cannot achieve horizontal scalability.
To the best of our knowledge, only the two systems described in~\Cref{sec:sharding} have experimentally demonstrated horizontal scalability.
However, these techniques rely on sharding; thus the performance highly depends on the transaction characteristics and the shard size parameter.
Furthermore, all the systems described in~\Cref{sec:taxonomy} are complete blockchain systems designed for a concrete application.
We wish to design a blockchain consensus protocol that can be used in many applications.
For example, the accounting of detailed internet traffic in Tribler~\cite{pouwelse2008tribler, multichain} and decentralised market in Tribler\footnote{\url{https://github.com/Tribler/tribler/blob/devel/Tribler/community/tradechain/__init__.py}}.
Hence our research question is the following.
\begin{displayquote}
\emph{How can we design a blockchain consensus protocol that is fault-tolerant,
horizontally scalable,
and able to reach global consensus?}
\end{displayquote}

To make sense of our research question, we must first explain \emph{blockchain consensus protocol}.
Typical blockchain systems are application specific.
For example, the application in Bitcoin (any many of its derivatives) is digital cash,
the application in Namecoin is domain name registration~\cite{namecoin},
and the application in Siacoin is cloud storage~\cite{siacoin}.
Underneath these applications lies the blockchain consensus protocol which has the goal of reaching consensus on transactions or state changes.
In the case of Bitcoin, it is PoW,
because it is not concerned with the structure or the semantics of the transactions.
In this work, we focus on the blockchain consensus protocol rather than any specific application because it is the scalability bottleneck and is necessary for any blockchain based application.
% Also, it should be possible to build any kind of application on top our system.
% We do not impose on a consensus algorithm, as long as it satisfies the properties of atomic broadcast which we describe in~\Cref{sec:overview-cons}.

In the remainder of this section, we expand on our research question and visit each aspect in detail.
At the end of this section, we describe some issues that we do not consider as a part of our design,
which is the result of not focusing on any particular application.

@@

\subsection*{Fault-tolerance}
% WHY WE NEED FAULT-TOLERANCE
Similar to many blockchain systems,
our system should be unaffected in the presence of powerful adversaries.
In particular, adversaries are Byzantine,
meaning that they can have arbitrary behaviour.
Thus anything is possible from simply omitting messages to colluding with each other to undermine the whole system.

There is another aspect of fault tolerance that is transaction verifiability.
All transactions must be eventually verifiable to maintain system integrity.
Further, the validation result must be consistent.
For example, if two different honest nodes attempt to verify the same transaction,
it cannot be the case that one node thinks that the transaction is valid, but the other thinks it is invalid.
Furthermore, just like transactions in cryptocurrencies, they should be tamper-proof and irrefutable.

@@

\subsection*{Limitations}
Since our consensus protocol is application neutral, it does not attempt to prevent the Sybil attack~\cite{douceur2002sybil}.
A good defence mechanism requires application specific information.
For example, a social network based Sybil defence mechanisms use random walks in real-world social graphs~\cite{yu2006sybilguard}.
Online marketplaces such as Amazon use the rating of buyer and sellers.
Cryptocurrencies use PoW to decide on the latest block.
% Since our system is application neutral, we do not provide an explicit Sybil defence mechanism.
For the same reason, we do not explicitly consider spam or denial of service attacks because,
without a concrete application,
it is hard to distinguish a super active user from an attacker.
For this work,
we assume that it is the role of the application developer to find a reputation mechanism or a trust model suitable for defending against such attacks.

### 2017-08-22 thesis/chapters/ch3-model.tex
Edit grammar etc.

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{architecture}
\caption{Overview of \textsc{Checo} architecture.}
\label{fig:architecture}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
    \centering
    \caption{An example of Extended TrustChain.
    The circles represent CP blocks,
    they also have hash pointers (arrow) but do not have references (dotted line).
    Note that the sequence number counter do not change, it is shared with TX blocks.
    CP blocks of index 0 (e.g. $c_{b, 0}$) are genesis blocks.}
    \label{fig:trustchain-bad-cp}
\end{figure}

@@

The communication complexity of Byzantine consensus algorithms typically grows polynomially w.r.t the number of nodes,
which prohibits us from running it on a large network.
For example, one of the most recent work tested with up to 104 nodes and the algorithm took almost 6 minutes to complete~\cite{miller2016honey}.
Thus, at the beginning of every Byzantine consensus algorithm execution,
we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm, and then run it.
After the algorithm completes,
the facilitators output a set of CP blocks which we call the consensus result.
The result is then propagated to the network and the next algorithm execution begins.

@@

\subsection*{The complete system}
The complete system is essentially the concurrent composition of the three aforementioned protocols,
all making use the Extended TrustChain data structure.

@@

% TODO big N does not follow our convention
In our model, we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ nodes are malicious and the inequality
$n \ge 3t + 1$ must hold.
This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol,
the exact reason becomes clear in~\Cref{sec:consensus-phase}.

@@

An improvement is use an epidemic protocol~\cite{eugster2004epidemic} instead, it is also known as gossip protocol.
Typical epidemic protocols work as follows.
Every node buffers every message it receives up to some buffer size $b$.
Then it forwards the messages $t$ number of times.
Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
The upside of using an epidemic protocol is that the communication cost is distributed more evenly between nodes,
which is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g., digest) to its $f$ random neighbours,
and only push the full message if the neighbour explicitly requests the message~\cite{leitao2007epidemic}.
With this, nodes typically only need to receive one consensus result message instead of $n$.

A down side of an epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
whereas broadcasting is constant time.
Another downside of some epidemic protocols (e.g., eager push) is that it is difficult to guarantee delivery.
It is especially true when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
If the delivery cannot be guaranteed, 
then we cannot guarantee liveness in our consensus protocol because a future facilitator may miss the memo.
Picking parameters are difficult in practice because the network configuration is unknown and the total number of nodes might also be unknown.

\subsection{Using timing assumption in the permissionless setting}
\label{sec:permissionless}
Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
However, in many applications it is often fine to make timing assumptions.
For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
A limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
In this section we show how making a timing assumption would allow us to operate in the permissionless setting.

At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-t$ \texttt{cp\_msg} messages,
which is the only place where we used $N$ as a parameter.
To introduce timing, instead of waiting for $N-t$ messages, we wait for some time $D$,
such that $D$ is sufficiently long for honest nodes to send their CP blocks to the facilitators.
Consequently, this removes the need for a PKI because the collected CP blocks may be from nodes that nobody has seen in the past.
However, choosing the parameter $D$ is difficult and depends on a number of factors such as the network condition, message size, and so on.
Overestimating it would make agreed fragments much longer than usual, which increases communication costs for validation.
Underestimating it would lead to unfairness where users with a poor internet connection will have a lower chance to be selected as a facilitator in the next round.
Nevertheless, there is a significant gain for making the timing assumption and that is the ability to operate in the permissionless setting which we explain next.

@@

\begin{figure}
    \includegraphics[width=0.7\textwidth]{compact}
    \centering
    \caption{The chain on the left represent direct chaining of full blocks,
    where the digest in ``Prev'' is simply the digest of the previous full block.
    The chain on the right uses compact blocks, represented by the smaller squares.
    Compact blocks also form a chain as before, but they each have a hash pointer to the full block, identified by ``Seq'' and ``Digest''.}
    \label{fig:compact}
\end{figure}

Concretely, we introduce an additional block type,
namely compact block.
Such blocks only have three attributes,
\begin{enumerate}
\item Seq---the sequence number of the corresponding full block,
\item Digest---the digest of the corresponding full block, and
\item Prev---the digest of the previous compact block.
\end{enumerate}
Each compact block has a corresponding full block (either a CP block or a TX block).
The relationship is uniquely identified with Seq or Digest.
Recall that our original validation protocol requires the nodes to send the agreed fragment containing the full blocks.
With compact blocks, it is only necessary to send the compact version of the agreed fragment.
The validation then proceeds in a similar fashion,
provided that the pair of the to-be-validated TX block is known.

@@

We offer two approaches to do total fork detection.
First and the easiest solution is to simply ask for the complete hash chain of the counterparty.
The verifier can be sure that there are no forks if the following conditions hold.
\begin{enumerate}
    \item The hash pointers are correct.
    \item All the CP blocks are in consensus.
    \item The TX of interest is in the chain.
\end{enumerate}
We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
Nodes employ caching to minimise communication costs, we call this effect spontaneous sharding.

The second approach is probabilistic but with only a constant communication overhead over our current design.
For a node, observe that if all of its agreed fragments has a transaction with an honest node,
then the complete chain is effectively validated in a distributed manner.
The only way for an attacker to make a fork is to make sure that the agreed fragment containing the fork has no transactions with honest nodes.
Such malicious behaviour is prevented probabilistically using a challenge-response protocol as follows.
Suppose node $A$ wish to make a transaction with node $B$.
$A$ first sends a challenge to $B$ asking it to prove that it holds a valid agreed fragment between some consensus round specified by $A$.
If $B$ provides a correct and timely response, then they run the transaction protocol as usual.
Otherwise, $A$ would refuse to make the transaction.

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{architecture}
\caption{The four components of \textsc{Checo}.}
\label{fig:architecture}
\end{figure}

@@

\begin{figure}
    \includegraphics[width=0.9\textwidth]{trustchain-bad-cp}
    \centering
    \caption{An example of Extended TrustChain.
    The circles represent CP blocks,
    they also have hash pointers (arrow) but do not have references (dotted line).
    Note that the sequence number counter does not change, it is shared with TX blocks.
    CP blocks of index 0 (e.g. $c_{b, 0}$) are genesis blocks.}
    \label{fig:trustchain-bad-cp}
\end{figure}

@@

The communication complexity of Byzantine consensus algorithms typically grows polynomially w.r.t the number of nodes,
which prohibits us from running it on a large network.
For example, one of the most recent works tested with up to 104 nodes and the algorithm took almost 6 minutes to complete~\cite{miller2016honey}.
Thus, at the beginning of every Byzantine consensus algorithm execution,
we randomly elect a set of nodes---called facilitators---to collect CP blocks from every other node and use those blocks as the input to the Byzantine consensus algorithm, and then run it.
After the algorithm completes,
the facilitators output a set of CP blocks which we call the consensus result.
The result is then propagated to the network,
and the next algorithm execution begins.

@@

\subsection*{The complete system}
The complete system is essentially the concurrent composition of the three protocols above;
all making use the Extended TrustChain data structure.

@@

% TODO big N does not follow our convention
In our model, we consider $N$ nodes, which is the population size.
$n$ of them are facilitators, $t$ nodes are malicious and the inequality
$n \ge 3t + 1$ must hold.
This inequality is from the work of Pease, Shostak and Lamport, where they show a network of $n$ nodes cannot reach Byzantine agreement with $t \ge n/3$~\cite{pease1980reaching}.
Further, the inequality $N \ge n + t$ must also hold so that there are enough honest nodes in the network to run the consensus protocol;
the exact reason becomes apparent in~\Cref{sec:consensus-phase}.

@@

An improvement is to use an epidemic protocol~\cite{eugster2004epidemic} instead, also known as gossip protocol.
Typical epidemic protocols work as follows.
Every node buffers every message it receives up to some buffer size $b$.
Then it forwards the messages $t$ number of times.
Every time the message is sent to $f$ random neighbours, $f$ is often called the fan-out.
The upside of using an epidemic protocol is that the communication cost is distributed more evenly between nodes.
It is especially true with a lazy push approach where the node who just received a message would push the message ID (e.g., digest) to its $f$ random neighbours,
and only push the full message if the neighbour explicitly requests the message~\cite{leitao2007epidemic}.
With this, nodes typically only need to receive one consensus result message instead of $n$.

A down side of an epidemic protocol is that it usually takes $O(\log N)$ time to infect the whole network,
whereas broadcasting is constant time.
Another downside of some epidemic protocols (e.g., eager push) is that it is difficult to guarantee delivery.
It is particularly the case when the parameters are not chosen correctly in a network that is only partially connected (but every node is nevertheless reachable).
If the delivery cannot be guaranteed, 
then we cannot guarantee liveness in our consensus protocol because a future facilitator may miss the memo.
Picking parameters are difficult in practice because the network configuration is unknown and the total number of nodes might also be unknown.

\subsection{Using timing assumption in the permissionless setting}
\label{sec:permissionless}
Our model is purely asynchronous, where we make no timing assumptions anywhere in the protocol.
However, in many applications, it is often fine to make timing assumptions.
For example, TCP relies on timeout for its retransmission and the \texttt{nLockTime} property in Bitcoin transactions makes the transaction unspendable until some time in the future (either Unix time or block height)~\cite{bitcoindevguide}.
A limitation of our system is that we use the parameter $N$ in our algorithms, which makes it unsuitable for the permissionless environment where users can join and leave at will.
In this section we show how making a timing assumption would allow us to operate in the permissionless setting.

At the start of our consensus phase (\Cref{sec:consensus-phase}), facilitators must wait for $N-t$ \texttt{cp\_msg} messages,
which is the only place where we used $N$ as a parameter.
To introduce timing, instead of waiting for $N-t$ messages, we wait for some time $D$,
such that $D$ is sufficiently long for honest nodes to send their CP blocks to the facilitators.
Consequently, this removes the need for a PKI because the collected CP blocks may be from nodes that nobody has seen in the past.
However, choosing the parameter $D$ is difficult and depends on many factors such as the network condition, message size, and so on.
Overestimating it would make agreed fragments much longer than usual, which increases communication costs for validation.
Underestimating it would lead to unfairness where users with a poor internet connection will have a lower chance to be selected as a facilitator in the next round.
Nevertheless, there is a significant gain for making the timing assumption,
and that is the ability to operate in the permissionless setting which we explain next.

@@

\begin{figure}
    \includegraphics[width=0.6\textwidth]{compact}
    \centering
    \caption{The chain on the left represents direct chaining of full blocks,
    where the digest in ``Prev'' is simply the digest of the previous full block.
    The chain on the right uses compact blocks, represented by the smaller squares.
    Compact blocks also form a chain as before, but they each have a hash pointer to the full block, identified by ``Seq'' and ``Digest''.}
    \label{fig:compact}
\end{figure}

Concretely, we introduce an additional block type,
namely compact block.
Such blocks only have three attributes,
\begin{enumerate}
\item Seq---the sequence number of the corresponding full block,
\item Digest---the digest of the corresponding full block, and
\item Prev---the digest of the previous compact block.
\end{enumerate}
Each compact block has a corresponding full block (either a CP block or a TX block).
The relationship is uniquely identified with Seq or Digest.
Recall that our original validation protocol requires the nodes to send the agreed fragment containing the full blocks.
With compact blocks, it is only necessary to send the compact version of the agreed fragment.
The validation then proceeds similarly,
provided that the pair of the to-be-validated TX block is known.

@@

We offer two approaches to do total fork detection.
First and the easiest solution is to ask for the complete hash chain of the counterparty.
The verifier can be sure that there are no forks if the following conditions hold.
\begin{enumerate}
    \item The hash pointers are correct.
    \item All the CP blocks are in consensus.
    \item The TX of interest is in the chain.
\end{enumerate}
We use this approach in our prior work on Implicit Consensus~\cite{implicitconsensus}.
Nodes employ caching to minimise communication costs, and we call this effect spontaneous sharding.

The second approach is probabilistic but with only a constant communication overhead over our current design.
For a node, observe that if all of its agreed fragments has a transaction with an honest node,
then the complete chain is effectively validated in a distributed manner.
The only way for an attacker to make a fork is to ensure that the agreed fragment containing the fork has no transactions with honest nodes.
Such malicious behaviour is prevented probabilistically using a challenge-response protocol as follows.
Suppose node $A$ wish to make a transaction with node $B$.
$A$ first sends a challenge to $B$ asking it to prove that it holds a valid agreed fragment between some consensus round specified by $A$.
If $B$ provides a correct and timely response, then they run the transaction protocol as usual.
Otherwise, $A$ would refuse to make the transaction.
